Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread

### 2017-07-28 paper.tex
minor edits

Adaptive introgression has three components: hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and selection against undesirable wild background.
To date, literature on crop-wild gene flow has focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
The history of natural introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
Preliminary results suggest a need to expand our conception of domestication to include the broadening of the genetic base of crops that occurred during post-domestication expansion through gene flow with newly encountered wild relatives.

@@

The relationship between maize (\emph{Zea mays} ssp. \emph{mays}) and the teosinte \emph{Zea mays} ssp. \emph{mexicana} (hereafter, \emph{mexicana}) offers a prime case study of adaptive wild-to-crop introgression.
Maize was domesticated from \emph{Zea mays} ssp. \emph{parviglumis} (hereafter, \emph{parviglumis})  approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
Introgression from \emph{mexicana} to maize in the Central Plateau has been reported based on both morphological \cite {wilkes1977} and molecular \cite{vanHeerwaarden2011, doebley1987} data.
However, \citep{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
%The following is not fully clear.  Cut or revise.
The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting using both HAPMIX and the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
These introgressed segments overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
Height differences were not detected under lowland conditions.
%Among the nine regions, three span the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
%In summary, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.

@@

Populations of \emph{mexicana} cannot be found outside of the highlands of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
%Several questions regarding \emph{mexicana} introgression yet remain.
%Has the introgression from \emph{mexicana} spread to multiple highland populations? 
%How do highland-adaptation traits differ among populations with and without introgression from \emph{mexicana}?
A recent study \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate whether maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
%On chromosome 4 (Figure \ref{fig:fd}), Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al. (2013) \citep{hufford2013}.
%The signal of introgression is absent for the other three populations.
%On chromosome 5, the signals of introgression (the peak region of the distribution) are present in MexHigh, GuaHigh and Southwestern US Highland (SW\_US), but not the other two.
%More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
Overall, analyses revealed that maize landraces with \emph{mexicana} introgression were transferred to nearby high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry. 
%The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects. 
%The stronger genetic drift introduced by the smaller historical population size increased homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
%The high frequency of deleterious alleles caused by stronger genetic drift in the Andes population, together with the reduced efficiency of selection against deleterious sites, contributes to the observed higher mutation load in the well-isolated maize population.
%Although it is clear that the absence of introgression from wild relatives provided fewer genetic resources for highland adaptation (making the highland adaptation in the Andes unique), it is yet unknown whether or not being out of reach of wild relatives is a reason for reduced fitness in the Andes population.
%Furthermore, the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
Historically, \emph{mexicana} haplotypes appear to have played an important role in adaptation of maize to challenging high-elevation conditions.
Breeding programs for highland maize may further benefit from drawing on \emph{mexicana} germplasm, particularly in regions like the Andes where \emph{mexicana} alleles are not known to have spread.

@@

%Domesticated and wild barley belong to the same species, \emph{Hordeum vulgare}, and are capable of producing viable offpspring via hybridization \cite{von1995ecographical}.
Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) was domesticated at least twice roughly 8,000 to 10,000 BP: once from the wild subsp. \emph{spontaneum} in the Fertile Crescent and once from subsp. \emph{spontaneum} var. \emph{agriocrithon} in Tibet \cite{takahashi1955origin, badr2000origin, azhaguvel2007phylogenetic, haberer2015barley, ren2013tibet, dai2012tibet}.
%\gmj{While morrell2007 cites the Zargos mountains as a center of domestication, this paper is cited by ren2013tibet, which states that the findings of morrell2007genetic are just that a second origin exists east of the Fertile Crescent. It is cited by dai2012tibet and the poets paper, but little is said of its mention of the zargos mountains.  I think that the mention of the Zargos mountains was just as a dividing line between the two spheres of influence between the two domestication centers.}
Presently, the distribution of subsp \emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
Barley/\emph{spontaneum} hybrids are fertile and found spontaneously when wild and domesticated barley co-occur.
Introgression between wild and domesticated barley is frequent \cite{dai2012tibet}, at times occurring over distances greater than a kilometer \cite{hillman2001new}.

@@

Poets and co-authors \cite{Poets2015} investigated the range-wide contribution of wild barley to landraces, assessing both genome-wide and geographical patterns.
This study identified several lines of evidence consistent with wild introgression aiding the dispersal and adaptation of domesticated barley.
Genomic regions of shared ancestry were detected linking individual landraces to a number of wild relative populations, suggesting landraces may have received wild introgression on a continual basis during post-domestication expansion.
However, barley landraces showed an excess of ancestry from nearby wild relatives, indicating a prevalence of local and potentially adaptive gene flow.
Low linkage disequilibrium and small blocks of identity by state indicated even these locally introgressed regions are old, perhaps dating back to the early expansion of barley following domestication.
While these results are suggestive, wild barley haplotypes have yet to be definitely linked to specific local adaptations in landraces.

@@

The story of Asian rice (\emph{Oryza sativa}) domestication is still debated, with hybridization between wild and domesticated rice contributing to the complexity of this crop's history.
On one hand, some genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from \emph{rufipogon}, with later divergence of \emph{indica} and \emph{japonica} \cite{molina2011molecular}.
Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia. %Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."

>>>>

Adaptive introgression has three components: hybridization between two genomes, backcrossing to one of the parents, and selection on different recombinant genotypes with progressively diminished linkage drag \cite{barton2001role, Feuillet200824}.
In domesticated species, adaptive introgression would consist of crop/wild hybrids backcrossing to a crop, retention and increase in frequency of adaptive wild haplotypes in the crop, and selection against undesirable wild background.
To date, literature on crop-wild gene flow has focused on the risk of transgene introgression from domesticated crops into wild relatives (for a review, \cite{stewart2003transgene}) and on modern plant breeding efforts to introgress desired traits from wild relatives (for a review, \cite{Dempewolf2017}).
The history of natural and potentially adaptive introgression of wild alleles into domesticated crops over evolutionary timescales has received considerably less attention.
However, new tools and methods have recently been employed to detect genome-wide patterns of introgression, granting new insights into the prevalence of adaptive introgression in crop histories.
Preliminary results suggest a need to expand our conception of domestication to include the broadening of the genetic base of crops that occurred during post-domestication expansion through gene flow with newly encountered wild relatives.

@@

The relationship between maize (\emph{Zea mays} ssp. \emph{mays}) and the teosinte \emph{Zea mays} ssp. \emph{mexicana} (hereafter, \emph{mexicana}) offers a prime case study of adaptive wild-to-crop introgression.
Maize was domesticated from \emph{Zea mays} ssp. \emph{parviglumis} (hereafter, \emph{parviglumis})  approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
Introgression from \emph{mexicana} to maize in the Central Plateau has been reported based on both morphological \cite {wilkes1977} and molecular \cite{vanHeerwaarden2011, doebley1987} data.
However, \citep{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
%The following is not fully clear.  Cut or revise.
The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting methods including HAPMIX and the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
These introgressed segments overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
Height differences were not detected under lowland conditions.
%Among the nine regions, three span the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
%In summary, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.

@@

Populations of \emph{mexicana} cannot be found outside of the highlands of Mexico, yet maize has colonized and adapted to high elevation in a number of other regions.
%Several questions regarding \emph{mexicana} introgression yet remain.
%Has the introgression from \emph{mexicana} spread to multiple highland populations? 
%How do highland-adaptation traits differ among populations with and without introgression from \emph{mexicana}?
A recent study \cite{wang2017} employed the ABBA-BABA and $\hat{f_{d}}$ statistics to evaluate whether maize with \emph{mexicana} introgression was transferred to other highland regions or whether highland adaptation was obtained \emph{de novo} outside of Mexico.
%On chromosome 4 (Figure \ref{fig:fd}), Mexican highland (MexHigh) and Guatamalan highland (GuaHigh) exhibited strong evidence of introgression from \emph{mexicana}, and the peaks of distribution corresponded to the region identified in Hufford et al. (2013) \citep{hufford2013}.
%The signal of introgression is absent for the other three populations.
%On chromosome 5, the signals of introgression (the peak region of the distribution) are present in MexHigh, GuaHigh and Southwestern US Highland (SW\_US), but not the other two.
%More details on the other chromosomes can be found in \cite{Wang2015manuscript}.
Overall, analyses revealed that  \emph{mexicana} introgressions were transferred to nearby high elevation regions in Guatemala and the southwestern United States, but more distant high elevation regions (\emph{e.g.,} the Andes) showed no \emph{mexicana} ancestry. 
%The Andean maize, the population totally isolated from the occurrence of any teosinte species, underwent the severest historical bottleneck, as a population in the front wave of the serial founder effects. 
%The stronger genetic drift introduced by the smaller historical population size increased homozygosity but reduced the heterozygosity of the population, and it also extended the length of homozygosity.
%The high frequency of deleterious alleles caused by stronger genetic drift in the Andes population, together with the reduced efficiency of selection against deleterious sites, contributes to the observed higher mutation load in the well-isolated maize population.
%Although it is clear that the absence of introgression from wild relatives provided fewer genetic resources for highland adaptation (making the highland adaptation in the Andes unique), it is yet unknown whether or not being out of reach of wild relatives is a reason for reduced fitness in the Andes population.
%Furthermore, the question of whether convergent evolution occurs between populations with and without introgression from wild relatives is a key topic for future studies in maize.
Since \emph{mexicana} haplotypes appear to have played an important role historically in adaptation of maize to challenging high-elevation conditions, modern breeding programs may further benefit from drawing on \emph{mexicana} germplasm, particularly in regions like the Andes where \emph{mexicana} alleles are not known to have spread.

@@

%Domesticated and wild barley belong to the same species, \emph{Hordeum vulgare}, and are capable of producing viable offpspring via hybridization \cite{von1995ecographical}.
Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) was domesticated at least twice roughly 8,000 to 10,000 BP: once from the wild subsp. \emph{spontaneum} in the Fertile Crescent and once from subsp. \emph{spontaneum} var. \emph{agriocrithon} in Tibet \cite{takahashi1955origin, badr2000origin, azhaguvel2007phylogenetic, haberer2015barley, ren2013tibet, dai2012tibet}.
%\gmj{While morrell2007 cites the Zargos mountains as a center of domestication, this paper is cited by ren2013tibet, which states that the findings of morrell2007genetic are just that a second origin exists east of the Fertile Crescent. It is cited by dai2012tibet and the poets paper, but little is said of its mention of the zargos mountains.  I think that the mention of the Zargos mountains was just as a dividing line between the two spheres of influence between the two domestication centers.}
Presently, the distribution of subsp \emph{spontaneum} stretches from the eastern Mediterranean through the Middle-East to west-central Asia spanning clines in temperature, precipitation, soil type, and altitude \cite{nevo2010drought}.
Barley-\emph{spontaneum} hybrids are fertile and found spontaneously when these taxa co-occur.
Introgression between wild and domesticated barley is frequent \cite{dai2012tibet}, at times occurring over distances greater than a kilometer \cite{hillman2001new}.

@@

Poets and co-authors \cite{Poets2015} recently investigated the range-wide contribution of wild barley to landraces, assessing both genome-wide and geographical patterns.
This study identified several lines of evidence consistent with wild introgression aiding the dispersal and adaptation of domesticated barley.
Genomic regions of shared ancestry were detected linking particular landraces to several distinct wild relative populations, suggesting landraces may have received wild introgression on a continual basis during post-domestication expansion.
However, barley landraces showed an excess of ancestry from nearby wild relatives, indicating a prevalence of local and potentially adaptive gene flow.
Low linkage disequilibrium and small blocks of identity by state indicated even these locally introgressed regions are old, perhaps dating back to the early expansion of barley following domestication.
While these results are suggestive, wild barley haplotypes have yet to be definitely linked to specific local adaptations in landraces.

@@

The story of Asian rice (\emph{Oryza sativa}) domestication is still debated, with hybridization between wild and domesticated rice contributing to the complexity of this crop's history.
Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{indica} and \emph{japonica} \cite{molina2011molecular}.
Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia. %Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."
